Mg Depleted and K Enriched Stars from LAMOST Spectral Data

### 2018-01-11 article/ms.tex
add ack list
Addressed many of the todos, begun construction of an abstract in Nature structure

% Abstract of the paper
\begin{abstract}
454180 RGB stars from the low resolution LAMOST DR2 survey were searched for anomalously high K and depleted Mg signatures. This study represents the largest search for field stars containing the Mg-K anti-correlation so far identified only in the globular NGC 2419 and NGC 2808. 23 stars displaying absorption spectra with over-abundant K and depleted Mg accross a range of metalicities were identified, but none were identified as having [Mg/Fe] or [K/Fe] consistent with the Mg depleted population found in NGC 2419. The abundances in these 23 stars are estimated at ranging between [INSERT RANGE FOR K AND MG HERE BASED OFF LAMOST DATA]. High resolution spectra for three stars (1 of which is not included among the final count of 23 spectral candidates for the signiature for reasons relating to the LAMOST signal quality for that star, despite having the highest K abundance of the three at [K/Fe]=1.54) selected based on observability characteristics were obtained using the Las Campanas Observatory, and for these stars elemental abundances for heavy elements such as Sc and V are also presented.
\end{abstract}

@@

NGC 2419 is the Milky Way's third most massive globular cluster, and its chemical composition makes it perhaps the most unusual star cluster in the Galaxy. Recent spectroscopic studies of red giant branch (RGB) stars in NGC 2419 revealed a strong anti-correlation between Mg and K abundances in nearly half of the studied stars, and weaker abundance relations in Si, Sc, Ca, Ti and V \citep{mucciarelli2012,cohenkirby2012}. Depletions of Mg within globular clusters are not uncommon, but are always accompanied with an increase in [Fe/H] \todo{[citations needed; possibly remove]}. One mechanism for this phenomena is the introduction of $\alpha$-poor and Fe-rich material released from Type Ia supernovae which locally pollutes the more `normal' [Mg/Fe] ratios that arise from Type II supernovae. However, the two populations in NGC 2419 are indistinguishable in their [Fe/H] abundance \citep{cohenkirby2012}. Furthermore, Type Ia supernovae enrichment cannot account for the enhanced K abundance or the similar  abundance trends observed in NGC 2419. Currently there is no known nuclear process that can explain the Mg-K abundance variations  without invoking unrealistic temperature conditions or by altering reaction rates by several orders of magnitude.

@@

A  targeted search looking [K/Fe] in RGB stars from clusters NGC 6752, NGC 6121, NGC 1904, and $\omega$ Centauri as well as 21 field stars \todo{21 field stars or 21 clusters?? I thought carretta looked at lots of clusters, no?} found in all cases K abundances falling within the bounds of the Mg normal population in NGC 2419 \citep{carretta2013}. \todo{The exception is NGC 2808......}
With the exception of NGC 2808, in which all 4 of its known Mg depleted stars have been found to have an anti-correlation with K \citep{mucciarelli2015}, this signature has thus far been found to be contained within NGC 2419. The fact that the Mg-K anti-correlation is confined to these very few specific globular clusters would seem to imply either a small population  of unusual polluter stars of some sort or a single extremely massive polluter star.

An early modelling attempt at replicating the anomalous Mg/K abundances assumed hot-bottom burning in AGB and super-AGB (SAGB) stars \citep{ventura2012}. This study succeeded in reproducing the Mg-K anti-correlation through the nuclear reaction pathway \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(e^-,\nu)^{37}Cl(p,\gamma)^{38}Ar(p,\gamma)^39K} \citep{iliadis2016} for the production of K. However, abundance patterns consistent with NGC 2419 were only achievable if the reaction cross section was set to 100 times the commonly accepted rate, or if the \todo{stellar interior?} temperatures could be increased to 150 MK \todo{from what?}.
% made higher than currently accepted values, achieving a burning temperature of 150 MK.

@@

It was found that (with the exception of extremely high densities > $10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^{-4}$ and 10$^8$ g/cm$^3$ were required. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential polluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was Novae, although the lack of detailed models of white dwarfs accreting material similar to NGC 2419 either from a binary companion or the intra cluster medium limit appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. However, based on current novae frequency in globular clusters determined by \cite{kato2013novae}, \cite{iliadis2016} conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

Another proposed  polluter candidate is a pair instability super nova \citep[PISN;][]{carretta2013}. Unique to extremely massive population III stars, these events involve the total destruction of the star, with no black hole remnant being left behind. The main argument for this idea is that the extreme rarity of these events, coupled with the huge masses of processed material involved, could potentially allow the entire signature in NGC 2419 to be the result of a single event, explaining why it is not seen in any other globular clusters. Further, the signature odd-even proton number abundance pattern associated with PISNs could perhaps explain the anti-correlation between Mg and K, as well as the enhancements in Sc and V observed by \cite{cohenkirby2012}. \todo{Do all the abundance variations like Sc, Si, Ca, Ti, and V go in the expected direction for PISN?}

@@

The data set of \LamostGiants\ \lamost\ giants was prepared by \citet{ho2017} for use with the supervised machine learning code \tc. The \lamost\ data was shifted to rest frame and continuum-normalised before it was resampled to a common uniform wavelength grid between 3905\,\AA\ and 9000\,\AA. \tc\ was used to infer $\teff$, $\logg$, [Fe/H], and [$\alpha$/Fe] using 9,952 stars in common between \lamost\ and the higher spectral resolution \apogee\ survey through the use of a data-driven spectral model. For details regarding the preparatory work, model generation, and label transfer between the \apogee\ and \lamost\ surveys, the reader is directed to \citet{ho2017}.

@@

We fitted a gaussian profile to the flux residuals for all three absorption lines in the Mg triplet (5167 \AA, 5172 \AA, 5184 \AA) as well as the K doublet (7665 \AA, 7699 \AA) for all \LamostGiants\ in \lamost\ data release 2 (DR2). For every star we recorded the profile amplitudes, wavelengths, widths, as well as associated measurement uncertainties of these quantities. We identified candidates by requiring they match at least one of the following three quality filters.
\begin{enumerate}
\item We required the amplitude of the profile at Mg 5184 to exceed $a_{{\rm Mg} @ 5184} > 0.05$ and the amplitude of the profile at K 7665 to exceed $a_{{\rm K}\,@\,7665} < 0.05$. Both amplitudes must be measured at more than $3\sigma$ ($|a|/\sigma_{a} \geq 3$).
\item We required two of the three Mg triplet lines to satisfy $a > 0.05$ and at least one K line to have $a < 0.05$, and for those amplitudes to have $|a|/\sigma_{a} \geq 3$.
\item We required two of the three Mg triplet lines to have $a > 0$ and both K lines to have $a < 0$, and for the spectra to have a signal-to-noise ($S/N$) ratio of $S/N > 30$ in \lamost, and a reported $\chi_{r}^2 < 3$ from \tc.
\end{enumerate}

@@

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

>>>>

% Abstract of the paper
\begin{abstract}

Stellar absorption spectra act as a fossil record of the environment of primordial conditions. Stars displaying unusual elemental abundances provide are of particular interest, as they offer the opportunity to learn about astrophysical events and objects that may no longer exist in our Galaxy. Stars with depleted Mg and enhanced K abundances present just such an opportunity. Thus far found only in two massive globular clusters NGC 2419 and NGC 2808, the primordial origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. A satisfactory explanation for this signature has potential ramifications on theories of globular cluster formation and the long-standing problem of the underestimation of even galaxy-normal K abundances. The apparent exclusivity of this Mg-K anti-correlation to globular clusters is the focus of this survey. Here we present 113 field stars showing depleted Mg and enhanced K, though none with abundances extreme enough to be comparable to the stars identified in NGC 2419.

%454180 RGB stars from the low resolution LAMOST DR2 survey were searched for anomalously high K and depleted Mg signatures. This study represents the largest search for field stars containing the Mg-K anti-correlation so far identified only in the globular NGC 2419 and NGC 2808. 23 stars displaying absorption spectra with over-abundant K and depleted Mg accross a range of metalicities were identified, but none were identified as having [Mg/Fe] or [K/Fe] consistent with the Mg depleted population found in NGC 2419. The abundances in these 23 stars are estimated at ranging between [INSERT RANGE FOR K AND MG HERE BASED OFF LAMOST DATA]. High resolution spectra for three stars (1 of which is not included among the final count of 23 spectral candidates for the signiature for reasons relating to the LAMOST signal quality for that star, despite having the highest K abundance of the three at [K/Fe]=1.54) selected based on observability characteristics were obtained using the Las Campanas Observatory, and for these stars elemental abundances for heavy elements such as Sc and V are also presented.
\end{abstract}

@@

NGC 2419 is the Milky Way's third most massive globular cluster, and its chemical composition makes it perhaps the most unusual star cluster in the Galaxy. Recent spectroscopic studies of red giant branch (RGB) stars in NGC 2419 revealed a strong anti-correlation between Mg and K abundances in nearly half of the studied stars, and weaker abundance relations in Si, Sc, Ca, Ti and V \citep{mucciarelli2012,cohenkirby2012}. Depletions of Mg within dwarf spheroidal (dSph) galaxies are not rare \citep{mucciarelli2012}, but are always accompanied with an increase in [Fe/H].
%\todo{[citations needed; possibly remove. RESOLVED, addapted statement slightly. relevant sources: Mucciarreli and Tsujimoto.]}
The mechanism for this Mg depletion is the introduction of $\alpha$-poor and Fe-rich material released from Type Ia supernovae \citep{tsujimoto2012first} which locally pollutes the more `normal' [Mg/Fe] ratios that arise from Type II supernovae. However, the two populations in NGC 2419 are indistinguishable in their [Fe/H] abundance \citep{cohenkirby2012}. Furthermore, Type Ia supernovae enrichment cannot account for the enhanced K abundance or the similar  abundance trends observed in NGC 2419. Currently there is no known nuclear process that can explain the Mg-K abundance variations  without invoking unrealistic temperature conditions or by altering reaction rates by several orders of magnitude.

@@

A  targeted search looking [K/Fe] in RGB stars from clusters NGC 6752, NGC 6121, NGC 1904, and $\omega$ Centauri as well as 21 field stars 
%\todo{21 field stars or 21 clusters?? I thought carretta looked at lots of clusters, no?} RESOLVED: Info accurate, no action taken.
 found in all cases K abundances falling within the bounds of the Mg normal population in NGC 2419 \citep{carretta2013}.
 % \todo{The exception is NGC 2808......} RESOLVED: Rephrased paragraph.
NGC 2808 is the only cluster other than NGC 2419 where the anticorrelation between Mg and K has been observed. All 4 of its known Mg depleted stars have been found to have an anti-correlation with K \citep{mucciarelli2015}, although the amplitude of these abundances was weaker than that in NGC 2419. The fact that the Mg-K anti-correlation is apparently confined to these very few globular clusters would seem to imply either a small population  of unusual polluter stars of some sort or a single extremely massive polluter star.

An early modelling attempt at replicating the anomalous Mg/K abundances assumed hot-bottom burning in AGB and super-AGB (SAGB) stars \citep{ventura2012}. This study succeeded in reproducing the Mg-K anti-correlation through the nuclear reaction pathway \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(e^-,\nu)^{37}Cl(p,\gamma)^{38}Ar(p,\gamma)^39K} (corrected as per \cite{iliadis2016}) \todo{Ask andy if its right to cite iliadis here, and why he removed the 'corrected as per' bit.} for the production of K. However, abundance patterns consistent with NGC 2419 were only achievable if the reaction cross section was set to 100 times the commonly accepted rate or if the temperature at the base of the envelope  temperatures could be made to exceed $\sim$ 150 K.

@@

It was found that (excepting the case of extremely high densities > $10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^{-4}$ and 10$^8$ g/cm$^3$ were required. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential polluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was novae, although the lack of detailed models of white dwarf accretion of material similar to that of NGC 2419 either from a binary companion or the intracluster medium limits appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. However, based on current novae frequency in globular clusters determined by \cite{kato2013novae}, \cite{iliadis2016} conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

Another proposed  polluter candidate is a pair-instability super nova (PISN) \citep[PISN;][]{carretta2013}. Unique to extremely massive Population III stars, these events involve the total destruction of the star, with no black hole remnant being left behind. The main argument for this idea is that the extreme rarity of these events, coupled with the huge masses of processed material involved, could potentially allow the entire signature in NGC 2419 to be the result of a single event, explaining why it is not seen in any other globular clusters. However, the signature odd-even proton number abundance pattern associated with PISNs is absent from NGC 2419 \citep{cohenkirby2012}, making this particular scenario unlikely. However, some other type of extremely massive star exploding in a single event would certainly go a long way to explaining the rarity of the signal.
% \todo{Do all the abundance variations like Sc, Si, Ca, Ti, and V go in the expected direction for PISN?} RESOLVED: You are quite correct, I was entirely mistaken. In fact, the odd-even abundance pattern is absent from NGC 2419 in general. segment is edited to reflect this.

@@

The data set of \LamostGiants\ \lamost\ giants was prepared by \citet{ho2017} for use with the supervised machine learning code \tc. The \lamost\ data was shifted to rest frame and continuum-normalised \todo{double check with andy. I could have sworn she specifically said that it WASNT continuum normalised...} before it was resampled to a common uniform wavelength grid between 3905\,\AA\ and 9000\,\AA. \tc\ was used to infer $\teff$, $\logg$, [Fe/H], and [$\alpha$/Fe] using 9,952 stars in common between \lamost\ and the higher spectral resolution \apogee\ survey through the use of a data-driven spectral model. For details regarding the preparatory work, model generation, and label transfer between the \apogee\ and \lamost\ surveys, the reader is directed to \citet{ho2017}.

@@

We fitted a gaussian profile to the flux residuals for all three absorption lines in the Mg triplet (5167 \AA, 5172 \AA, 5184 \AA) as well as the K doublet (7665 \AA, 7699 \AA) for all \LamostGiants\ giants in \lamost\ data release 2 (DR2). For every star we recorded the profile amplitudes, wavelengths, widths, as well as associated measurement uncertainties of these quantities. We identified candidates by requiring they match at least one of the following three quality filters.
\begin{enumerate}
\item We required the amplitude $A$ of the profile at the Mg 5184 \AA \ line to exceed $A_{{\rm Mg} @ 5184} > 0.05$ and the amplitude of the profile at the K 7665 line to exceed $A_{{\rm K}\,@\,7665} < 0.05$. Both amplitudes must also be measured at more than $3\sigma$ ($|A|/\sigma_{A} \geq 3$).
\item We required two of the three Mg triplet lines to satisfy $A > 0.05$ and at least one K line to have $A < 0.05$, and for those amplitudes to have $|A|/\sigma_{A} \geq 3$.
\item We required two of the three Mg triplet lines to have $A > 0$ and both K lines to have $A < 0$, and for the spectra to have a signal-to-noise ($S/N$) ratio of $S/N > 30$ in \lamost, and a reported $\chi_{r}^2 < 3$ from \tc.
\end{enumerate}

@@

- LAMOST
- ARC DP (Casey)
- Magellan/MIKE and Australian government
- others???
